---
title: "copy_landcover-focal"
author: "Miles Van Denburg"
date: "4/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


An overview of how to do a focal window to get land cover neighborhood statistics around each pixel (to mimic buffering)

## Step 1: Read in, crop, and rescale landcover data

Data are in 0-255.  Predictions probably shouldn't be made to far outside of the range of the extent of the samples, so crop down land cover data first (also makes for less compute)
```{r, eval=FALSE}
library(raster)
library(dplyr)
library(sf)
ma <- st_as_sf(maps::map("state", plot = FALSE, fill = TRUE)) %>% 
  filter(ID == "massachusetts")


# read in sample points
# setwd(dir = "../naturewave/extdata/sample_sites/")
getwd()
setwd("~/Spatial_Analysis_R/naturewave/extdata/sample_sites")
pts <- read_sf("sample_sites_projected.shp")

# read in landcover data
f <- dir("external/data", full.names = TRUE, pattern = ".tif")
landcover <- lapply(unname(f), function(x) {
  r <- raster(x)
  crs(r) <- crs(pts)
  r
})
names(landcover) <- gsub(".tif", "", basename(f))


# define slightly wider extent than that of the points, and set up mask from 
# state boundaries of MA to remove 0 value areas outside of state
e <- extend(extent(pts), 10000, -10000, 0, 0)
ma_crop <- ma %>% st_transform(crs = st_crs(pts)) %>% st_crop(., e)

# apply crop, mask and rescale to each layer
lc_crop <- lapply(landcover, function(x){  # x <- landcover[[1]]
  r <- crop(x, e)  # crop to box arpund pts
  r <- mask(r, ma_crop)  # mask out no data areas
  (r - cellStats(r, min)) / 255
})
lc_stack <- stack(lc_crop)  # stack layers

# quick look
plot(lc_stack[[1]])
plot(st_geometry(pts), add = TRUE)
```

## Step 2: run focal mean

Run focal mean, writing out focal results to disk to limit size in memory
```{r, eval = FALSE}
# function will give you a weights matrix for the desired neighborhood size, 
# given width (diam) and resolution of raster (r)
wmatrix <- function(diam, r) {
  d <- round(diam / r, 0)  # set window dimensions
  m <- matrix(nrow = d, ncol = d)  # make matrix
  m[1:length(m)] <- rep(1 / length(m), length(m))  # populate with weights
  m
}

# run focal mean, here for 2000 m
diam <- 2000  # buffer size (in meters)
w <- wmatrix(diam, res(landcover_stack)[1])  # weights matrix
lc2000 <- lapply(1:4, function(x) {  # x <- 1
  f <- paste0("external/data/", names(landcover_stack)[x], "_", diam, ".tif")
  r <- focal(landcover_stack[[x]], w, filename = f)
  return(r)
})
lc2000_stack <- stack(lc2000)  # stack

# quick look
plot(lc2000_stack)
```

The above is an example for how you can do the focal mean for a 2000 m window. You can repeat the above for different buffer sizes by copying the code and simply changing the `list` and `stack` object names and `diam`.

## Step 3: extract focal values to get predictors for points
```{r, eval = FALSE}
library(ggplot2)
pts <- cbind(pts, extract(lc_stack, pts))  # original lc
pts <- cbind(pts, extract(lc_2000s, pts))  # buffer points

ggplot(pts) + geom_point(aes(connect, connect_2000)) + 
  geom_abline(slope = 1) + xlim(0, 1) + ylim(0, 1)
ggplot(pts) + geom_point(aes(imperv, imperv_2000)) + 
  geom_abline(slope = 1) + xlim(0, 1) + ylim(0, 1)
ggplot(pts) + geom_point(aes(structure, structure_2000)) + 
  geom_abline(slope = 1) + xlim(0, 1) + ylim(0, 1)
ggplot(pts) + geom_point(aes(traffic, traffic_2000)) + 
  geom_abline(slope = 1) + xlim(0, 1) + ylim(0, 1)
```

